---
title: "Messy affair of DATEing"
description: "What to do when you dealing with DATES"
date: 2023-10-13
tag: ["r","r-tutorial"]
output: 
   rmarkdown::html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: hide
    code_downlod: true
    highlight: textmate


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dates are always a messy affair

My apologies if you thought I would be talking about dates, yes definitely I will be talking about dates but not that type of date you thinking but the actual *dates*.

Do you love dates, personally I do but they can be a mouthful of headaches if you understand what I mean. So here goes my dating scenes.

# Planning for dates

You cannot talk about *date* without specifying *time*. They are kind of intertwined somehow. In fact there is even an *ISO Standard 1860* specifically for this [here](https://www.iso.org/iso-8601-date-and-time-format.html), we should all follow this when dating and timing. Actual format *YYYY-MM-DD HH:MM:SS* or `r lubridate::now(tzone = "")`. 

When planning dates the details are really important.

# Preparing for dates

I would advocate using `lubridate` r package for easier understanding and manipulation of dates rather. I find it easier than base R but to each their own. Your date is your date and we cannot be all on the same date.

# Functions within dates


Function | Description 
-----| --------- 
`now()` | Returns date and time right now, at this instance
`today()` | Returns the date today
`parse_date_time` | Returns date from list of text you can specify the ordering e.g. `ymd, dmy, mdy` where y (year), m (month) and d (day) 


# Manipulating your date

You can always apply mathematical parameters when dealing with dates mainly `+, -`. Adding and subtracting days to come up with future or past dates.

With this we need to understand two principles:

1. `period` :- Human time value e.g. months, days, years, hours. R respective codes `days`, `hours`, `dweeks`.
1. `duration` :- Time value based in seconds calculating like a stop watch time for 86400 secs. Doesn't adhere to leap years, day light saving and all. R respective codes `ddays`, `dhours`, `dweeks` addition of prefix `d`.



When adhering to date rules i.e leap years we can use `%m+%` to add days and `%m-%` for subtracting.

```{r message=FALSE, warning=FALSE}

library(lubridate)

first_corona_case_date = parse_date_time("12 January 2020", orders = "dmy" ,tz = "")

first_corona_case_date


```
The first case of corona (COVID-19) was reported on `r first_corona_case_date` and 3 months down we first the a million case.

```{r message=FALSE, warning=FALSE}

million_corona_case_date = first_corona_case_date + months(3)

million_corona_case_date


```


The first million case reported was `r million_corona_case_date`


# How do we sort long distance dating

You planned for the date but your partner is in a different timezone how do you align the time. Now the particulars of your date apart from time has an extra component of timezone.

Assuming you are based in _Tokyo_ and your partner is abased in _Nariobi_ . First when using R we need to pick the list of city time zone `OlsonNames` e.g. `r OlsonNames()[1]`  for _Tokyo_ name is `r OlsonNames()[44]` and _Nairobi_ is `r OlsonNames()[322]`


When _Nairobi_ partner says the call is happening 8PM the parner in _Tokyo_ has to use the `with_tz()` function in R to calculate the time.

```{r message=FALSE, warning=FALSE}

Nariobi_time <- force_tz(as_datetime("2023-09-01 20:00:00"), tz = "Africa/Nairobi")
Nariobi_time

```

```{r message=FALSE, warning=FALSE}

Tokyo_time <- with_tz(Nariobi_time, tzone = "Asia/Tokyo")
Tokyo_time

```


# The date continues next time 
# When we schedule the next date hope I wasn't a boring partner.